### 2018-02-07 doc/ancestralPCA.tex
Add important existing files

\documentclass{article}[12pt]

\usepackage{epsfig} \usepackage[margin=1.5in]{geometry}
\usepackage[semicolon,authoryear]{natbib} \bibliographystyle{natbib}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{cleveref}
\usepackage{verbatim}

\newcommand{\bO}{\mathcal{O}}

\title{Computational Methods for Principle Component Analysis Applied to Ancestry Estimation}

\author{David Zhang}

\begin{document}

\maketitle

\section{Principle Component Analysis (PCA)}

\subsection{What is PCA}

Let there be $n$ individuals and $p$ features.
Moreover, let these points be centered.
Then we have $n$ points in a $p$-dimensional space,
with the origin being the center of mass of these points.
For any direction, we can project the points to it and find the variance of the projected points.
We are interested in the direction of the greatest variance.
This direction is the linear combination of the features that distinguishes the individuals the most,
which we call the first principle component (PC1).
We repeat the process in the quotient space induced by PC1 to find the second principle component (PC2) and so forth.
By using the first $k$ PC's, we can reduce the $p$-dimensional data to $k$-dimensional at the least cost of losing information.

\subsection{How to find the PCs}
To find the PC's, we record the information of this group in a $p \times n$ matrix $X$ and conduct singular value decomposition (SVD) to it.
SVD gives us the spectrum decomposition $X = UDV^T$, where $U$ is $p \times n$, $D$ and $V$ are $n \times n$.
Here $U$ and $V$ are unitary matrices,
and their columns are the eigenvectors of $XX^T$ and $X^T X$, respectively.
And $D$ is a diagonal matrix,
whose diagonal entries are the square roots of the nonzero eigenvalues of $X X^T$ and $X^T X$.
In other words, $X X^T = U D^2 U^T$ and $X^T X = V D^2 V^T$.
For PCA, the columns of $U$ are the first $n$ PC's,
the diagonal entries of $D$ are the standard deviations of the $n$ points projected to the $n$ PC's,
and the rows of $V$ are the $n$ points' standardized PC scores on the $n$ PC's,
which can be seen as $n$ points' coordinates in the space spanned by the $n$ PC's.

For more information on PCA, see \cite{jolliffe}.

\section{Application of PCA to Genotyping Data}

\subsection{The format of genotyping data}

PCA can be applied to genotyping data.
In particular, it is used to analyze single nucleotide polymorphisms (SNPs).
A SNP is a variation in a single nucleotide.
For example, most individuals may have an A base at a specific base position, but a minority of individuals may have a C base at this site.
Here we say that there is a SNP at this base position,
and A and C are said to be its major and minor alleles, respectively.
Since chromosomes come in pairs,
at each base position the minor allele may appear zero, one, or two times.
Thus by looking at certain base positions,
we receive a data of the form of a string of $0$, $1$, and $2$,
of length as large as hundreds of thousands.

\subsection{The usefulness of PCA on genotyping data}

The format of genotyping data makes it suitable for PCA.
A group of individuals is first selected as the reference group.
A testing individual is combined with the reference individuals to find its PC scores.
By using the first few PCs (often times four PCs are informative enough), 
we are able to convert the data from the form $3^p$ with $p \approx 10^6$ to $\mathbb{R}^k$ with $k \approx 4$,
which significantly reduces the complicatedness of the data and make it more intuitive to comprehend.
For example, we can find how similar is the testing individual to any of the reference individuals simply by eyeballing their PC1 vs PC2 plot.

\subsection{Ancestry estimation by PCA}

By applying PCA to genotyping data, we can retrieve information about many different aspects of an individual.
In this research project, we focus on using PCA to estimate an individual's ancestry history from his or her genotyping data.
Ancestry history is often a confounder for the association between certain gene and the trait being studied, such as having typej-II diabetes.
By estimating an individual's ancestry history,
researchers can eliminate the bias in the association between the gene and the trait caused by ancestry differences.
It has been found that individuals with the same ancestry history score similarly on the first four PCs of their genotyping data (see \cite{wang}).
Thus PCA provides a way to convert a person's ancestry history into real numbers,
which is useful for numerically estimating the previously mentioned confounding effect induced by ancestry history.

\section{Computational Methods for PCA}

There are at least two methods for finding the testing individual's PC scores.
Here we call them the ``once-for-everyone'' method and the ``once-for-all'' method.

\subsection{The ``once-for-everyone'' method}

In this method, we append the testing individual's genotyping data to the reference group's genotype matrix and compute the PC scores of the amended matrix.
This is the most standard and reliable computational method for PCA.
However, this method is relatively slow in computation.
Here all the reference individuals' PC scores must be recomputed when a new testing individual arrives.

\subsubsection{Standard implementation}

One of such implementation is the TRACE program by Wang et al.
An important addition in this program to the description above is Procrustes analysis.
Given two sets of points that are of the same number but live in spaces of different dimensions,
Procrustes analysis helps us to find a projection
that transforms the higher-dimensional set to the lower-dimensional set
such that the distance between the points after the transformation is the least.
This transformation improves the accuracy of the estimated ancestry.

The detailed algorithm of TRACE runs as follows.
Let $X$ be the $p \times n$ reference genotyping data matrix where each column represents an individual.
Moreover, let $y$ be the $p \times 1$ matrix containing the testing individual's genotyping data.
\begin{enumerate}
\item Compute $X^T X$.
  (Computational complexity: $\bO[n^2p]$.)  
\item Apply eigendecomposition to $X^T X$ to get $X^T X = V D^2 V^T$ (columns of $D$ and $V$ are sorted by PC variance with the first column corresponding to the greatest PC variance).
  ($\bO[n^3]$.)
\item Append $X^T y$ and its transpose to the last column and row of $X^T X$, respectively.
  The entry of the bottom-right corner is $y^T y$.
  This is exactly the same as computing $X'^T X'$,
  where $X' = (X, y)$ is the reference group with the testing individual added.
  ($\bO[np]$.)
\item Apply eigendecomposition to $X'^T X'$ to get $X'^T X' = V' D'^2 V'^T$.
  ($\bO[n^3]$.)
\item Project the first $k_2$ columns of $V'$ to the first $k_1$ columns of $V$ by using Procrustes analysis to get the transformed PC scores of the testing individual (the last row of $V$),
  where $k_1 < k_2 << n$.
  (Computational complexity is ignorable if $k_2$ is not large compared to $n$.)
  \item Go to step 3 for the next testing individual.
\end{enumerate}

This is a simplified version of the TRACE algorithm.
Some minor adjustments in the program are not mentioned above.
TRACE actually uses the unstandardized PC scores $VD$ rather than the standardized PC scores $V$.
The two treatments of the PC scores have the same computational complexity and the results are equivalent.
Moreover, TRACE has a slightly more sophisticated treatment of missing data.
In this report, we assume that no missing data exists to simply the comparison of the three methods.

By adding up the computational complexity of each step, we can see that the total computational complexity of TRACE is $\bO[n^2p + m(np + n^3)]$,
where $m$ is the number of testing individuals.

For more information about TRACE,
see \cite{wang} and TRACE's manual.

\subsubsection{Improvments}

The algorithm above can be speeded up by using a different way to calculate the PC scores, that is, matrix $V'$.
Notice that TRACE only uses the first $k_2$ columns of $V'$.
Thus we take advantage of this and use a method called online SVD that gives a close approximation of only some of the columns of $V'$ and saves computational complexity.
The detailed algorithm is:

\begin{enumerate}
\item Compute $X^T X$.
  (Computational complexity: $\bO[n^2p]$.)  
\item Apply eigendecomposition to $X^T X$ to get $X^T X = V D^2 V^T$.
  ($\bO[n^3]$.)
\item Calculate $U_3 = X V_3 D_3^{-1}$,
  where $V_3$ is the first $k_3$ columns of $V$,
  and $D_3$ is the diagonal matrix of the first $k_3$ diagonal entries of $D$.
  ($\bO[k_3 n p]$.)
\item Calculate 
  \[
    L = U_3^T y \quad \text{and} \quad K = y^T H,
  \]
  where $H$ is the normalized  $y - U_3L$
  ($\bO[k_3 p]$.)
\item Calculate $Q^T Q$, where
  \[
    Q = 
    \begin{bmatrix}
      D_3 & L \\
      0 & K
    \end{bmatrix}.
  \]
  ($\bO[k_3^3]$.)
\item Apply eigendecomosition to $Q^T Q$ to get $Q^T Q = V''_3 D''^2_3 V''^T_3$.
  ($\bO[k_3^3]$.)
\item Calculate
  \[
    V'_3 =
    \begin{bmatrix}
      V_3 & 0 \\
      0 & 1
    \end{bmatrix}
    V''_3.
  \]
  ($\bO[nk_3^2]$.)
\item Project the first $k_2$ columns of $V'_3$ to the first $k_1$ columns of $V$ by using Procrustes analysis to get the transformed PC scores of the testing individual,
  where $k_1 < k_2  < k_3 << n$.
  (Computational complexity is ignorable if $k_2$ is not large compared to $n$.)
  \item Go to step 4 for the next testing individual.
\end{enumerate}

The total computational complexity for $m$ testing individuals is $\bO[n^2p + n^3 + k_3np + m(k_3p + k_3^3 + nk_3^2])$, which is approximately $\bO[n^2p + m(k_3p + k_3^2n)]$ since $k_3 << n << p$.
This is a big save on the runtime compared to the TRACE algorithm,
if $k_3$ is small compared to $n$.

For more details of the online SVD method, see \cite{brand}.

\subsection{The ``once-for-all'' method}

In the ``once-for-everyone'' method,
we have to decompose the data matrix of both the reference group and the testing individual for every single testing individual.
Alternatively, we can find the PCs of the reference group and project the testing individual's genotyping data to that direction.
This way, eigendecomposition or SVD is done only once,
which significantly reduces computational complexity.
However, this method tends to underestimate the magnitude of the testing individual's PC scores when $n << p$,
which is often the case for genotyping data,
as discussed in \cite{dey}.
This is called the ``shrinking bias''.
This problem has been dealt with by the R package HDPCA by Dey.
The algorithm runs as follows:

\begin{enumerate}
\item Compute $X^T X$.
  (Computational complexity: $\bO[n^2p]$.)  
\item Apply eigendecomposition to $X^T X$ to get $X^T X = V D^2 V^T$.
  ($\bO[n^3]$.)
\item Calculate $W = (Y^T X) V D^{-1}$, where $Y$ is the data matrix for all the testing individuals. ($\bO[mpn + mn^2]$.)
  \item Input $D$ and $W$ to the pc\_adjust function in the HDPCA function to obtain the PC scores adjusted for the shrinking bias. ($\bO[p]$ for finding the number of distant spikes (function select.nspike), $\bO[n]$ for adjusting for the shrinking bias (function hdpc\_est))
\end{enumerate}

Then the total computational complexity is $\bO[n^2p + mn(p + n) + p + n]$,
which is approximately equal to $\bO[n^2p + mn(p+n)]$,
given that $m$ or $n$ is large.
The comparison of computational complexity of the three methods is shown in \Cref{tbl:cplx}.

\begin{table} 
  \centering
  \begin{tabular}{|l|l|}
    \hline
    Method & Computational Complexity \\ 
    \hline
    TRACE & $\bO[n^2 p + mn(p + n^2)]$ \\
    \hline
    TRACE + online SVD & $\bO[n^2 p + mk_3(p + k_3 n)]$ \\
    \hline
    HDPCA & $\bO[n^2p + mn(p+n)]$ \\
    \hline
  \end{tabular}
  \caption{
    Comparison of computational complexity for the three methods.
    Here $p$ is the number of SNPs,
    $n$ is the number of individuals in the reference group,
    and $m$ is the number of individuals in the  testing group.
    For Online TRACE, $k_3$ is the number of PCs calculated by online SVD, which is usually much less than $n$.}
  \label{tbl:cplx}
\end{table}

\section{Runtime Comparison}

We now do an empirical test to compare the efficiency of the three methods.

\subsection{Dataset}

We use the datasets from the Human Genome Diversity Project (HGDP) and the 1000 Genomes Project (1000GP).
The HGDP dataset has genotyping data about 600k SNPs on 900 individuals with different ancestry history.
The 1000GP dataset contains most of the SNPs in the HGDP dataset as well as many others.
Its sample size is about 2300.

For our test, we use the 1000GP dataset but only the SNPs that also appear in the HGDP dataset,
which are as many as about 570k.
Moreover, we remove the individuals who has at least one parent that is already in the dataset.

\subsection{Data Format}

We use TRACE's geno format.
Each row contains the genotype of an individual.
The first two columns are the individual's ID and ancestral group.
Starting at the third column, each column represents a SNP and is either 0, 1, or 2, which is the number of mutations, or -9, which is the value for missing data.
Each SNP's reading is centered and normalized before applying the PCA methods,
with the missing data replaced by the mean.

The vcf format is also acceptable, since TRACE provides the vcf2geno program that converts vcf files to geno files.

\subsection{Setting}

\section{Result}

We use the 1000Genomes dataset for testing the performance of the three methods.
Only the SNP's that have a minor allele frequency greater than 0.05 and appear in the HGDP dataset are selected.
Individuals whose parents are in the dataset are removed.
This filtering leaves us 570859 SNP's and 2481 individuals.
For the testing group, we use the 100 individuals at the bottom of the dataset.
For the reference group, we use the $n$ individuals at the top of the dataset,
where $n$ is set to 500, 1000, 1500, 2000, and 2300 to compare the runtimes.

\subsection{Estimations}

The PC scores calculated by the three methods are almost identical,
with the only exception of the PC3 and PC4 scores calculated by HDPCA for $n = 500$ being slightly different from those calculated by the other two methods.
(See the pdf files in the Google Drive folder for the plots.)

\subsection{Runtimes}

The three methods' runtimes for the learning phase are not different from each other by much.
However, for the analysis phase, Online PCA and HDPCA are significantly more efficient as $n$ increases.
This empirical result is consistent with the theoretical computational complexity.
In \Cref{tbl:runLearn} and \Cref{tbl:runAna},
we show the runtimes for different reference sizes.
We break the runtimes into two phases: learning and analysis.
The learning phase is required for only once,
and the analysis phase is required for every testing individual.

\begin{table} 
  \centering
  \begin{tabular}{|l|l|l|l|l|l|}
    \hline
    nTrain & 500 & 1000 & 1500 & 2000 & 2300 \\ 
    \hline
    TRACE & 138 & 482 & 1022 & 62557 & 85913 \\
    \hline
    TRACE + online SVD & 152 & 505 & 1407 & 36824 & 79027 \\
    \hline
    HDPCA & 139 & 477 & 1368 & 35825 & 76700 \\
    \hline
  \end{tabular}
  \caption{Empirical runtimes (sec) for the learning phase for different reference sizes}
  \label{tbl:runLearn}
\end{table}

\begin{table} 
  \centering
  \begin{tabular}{|l|l|l|l|l|l|}
    \hline
    nTrain & 500 & 1000 & 1500 & 2000 & 2300 \\ 
    \hline
    TRACE & 59 & 247 & 729 & 8531 & 7851 \\
    \hline
    TRACE + online SVD & 135 & 131 & 132 & 138 & 140 \\
    \hline
    HDPCA & 79 & 119 & 194 & 390 & 463 \\
    \hline
  \end{tabular}
  \caption{Empirical runtimes (sec) for the analysis phase for different reference sizes}
  \label{tbl:runAna}
\end{table}

\section{Discussion}

\begin{comment}

\section{Old}

Alternatively, we can find the PCs (directions of the greatest variances) of the reference group,
and then project the testing individual's data to them to find the PC scores.
The advantage of this method is that the PCs for the reference group only need to be calculated once for all testing individuals.
This method, though fast, tends to underestimate the magnitude of the PC scores when compared to the results given by the first method.
This ``shrinking bias'' is discussed in \cite{dey}.

This research project studies how either method can be improved.
The first method can be speeded up by alternating the SVD algorithm,
so that the computational complexity can be reduced at little cost of the accuracy of the result.
For the second method, the accuracy of the estimation can be improved by calculating and adjusting for the shrinking bias.
In this report, we compare the accuracy and runtimes of different ways to improve the two methods.
\subsection{TRACE}

TRACE has the most conceptually simple algorithm among the three methods.
Let $X$ be the $p \times n$ matrix ($n << p$) that contains the reference genotype,
where each row corresponds to a SNP and each column corresponds to an individual.
Moreover, let $y$ be the $p \times 1$ matrix of the genotype of the testing individual. The algorithm runs as follows:

\subsection{Online PCA}

The Online PCA method has the same algorithm except for calculating the testing individual's PC scores.
Notice that in TRACE, only the first $k_2$ PC's (that is, the first $k_2$ columns of $V'$) are used for the testing individual.
The Online PCA method can compute for the first $k$ PC scores of a data matrix and skip the rest.
The result is an approximation, but the runtime is significantly reduced.

More specifically, Online PCA takes the following steps:

\subsection{HDPCA}

A simple way to compute the PC scores for the testing individual is $w^T = y^T U$.
However, the resulted PC scores tend to be smaller than the PC scores obtained from eigendecomposing $X'^T X$ where $X^T = (X, y)$.
This shrinking bias can be adjusted by the HDPCA package in R.
The algorithm is as follows:

\end{comment}

\newpage

\begin{thebibliography}{}

\bibitem[Brand(2002)]{brand} Brand, M. (2002). Incremental singular value decomposition of uncertain data with missing values. {\it Computer Vision--ECCV 2002}, 707-720.

\bibitem[Dey \& Lee(2016)]{dey} Dey, R., \& Lee, S. (2016). Asymptotic properties of Principal Component Analysis and shrinkage-bias adjustment under the Generalized Spiked Population model. {\it arXiv preprint arXiv:1607.08647}.

\bibitem[Jolliffe(2002)]{jolliffe} Jolliffe, I. (2002). {\it Principal component analysis}. John Wiley \& Sons, Ltd.
Chicago

 \bibitem[Wang et al.(2015)]{wang} Wang, C., Zhan, X., Liang, L., Abecasis, G. R., \& Lin, X. (2015). Improved ancestry estimation for both genotyping and sequencing data using projection procrustes analysis and genotype imputation. {\it The American Journal of Human Genetics, 96}(6), 926-937.

\end{thebibliography}

\end{document}