### 2018-02-07 doc/paper.tex

\documentclass{article}[12pt]

\usepackage{epsfig} \usepackage[margin=1.5in]{geometry}
\usepackage[semicolon,authoryear]{natbib} \bibliographystyle{natbib}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{cleveref}
\usepackage{verbatim}
\usepackage{graphicx}

\newcommand{\bO}{\mathcal{O}}
\newcommand{\argmin}[1]{\underset{#1}{\operatorname{arg\;min}}\;}

\title{Principle Component Analysis Computational Methods
  for Estimating Ancestral Histroy from Genotype Data}

\begin{document}

\maketitle

\section{Introduction}

\section{Method}

\subsection{Model}

\begin{enumerate}
\item Data matrices:\\
  The data matrix contains the genotypes for the individuals.
  Each row represents a SNP, and each column represents an individual.
  Each entry is $0$, $1$, or $2$.
  The genotypes of the reference group are stored in the $p \times n$ matrix $X$,
  and those of the study group are stored in the $p \times m$ matrix $Y$.
  For most genetic data, $n \ll p$ and $m \ll p$.
\item Principle component analysis \\
  Given a multi-dimensional data, principle component analysis (PCA) looks for the directions, which are called the principle components (PC), in which the data has the greatest variance.
  The PC's form an orthonormal basis for the sample space,
  and we can ignore all but the first few PC's to reduce the dimension of the data.
  In particular, let $X=UDV^T$ be the singular value decomposition (SVD) of $X$,
  where $U$ is a $p \times n$ matrix whose columns are the PC's,
  $D$ is an $n \times n$ diagonal matrix storing the standard deviations of the PC's,
  and $V$ an $n \times n$ matrix such that each row contains the standardized PC scores of an individual.
  (Alternatively, we can also use eigendecomposition to find the matrices since $X^T X = VD^2V^T$ and $X X^T = UD^2U^T$.)
  We use $D$ to sort out the most influential $k$ PC's ($k \ll n$) and look at the coresponding columns of $V$ to find their PC scores.
  For estimating ancestral histroy from genotypes, often the first $4$ PC's will be sufficient for providing the information we need.

  PCA allows us to use the reference data to find the PC scores of the study data.
  The standard method is to append a study individual's genotype to the reference matrix and then conduct a SVD of the augmented matrix.
  The corresponding row in the matrix $V$ is the study individual's PC scores.
\end{enumerate}

\subsection{Computational Methods}

The usefulness of PCA for genetic data largely depends on its computational cost.
The computational complexity of SVD for a $p \times n+1$ matrix is $\min((n+1)^2p, (n+1)p^2) = n^2p $ for $1 \ll n \ll p$.
Thus the total computational complexity for performing PCA on $m$ individuals by using the standard method is $mn^2p$.
This value increases quadratically as we increase the reference sample size to increase the accuracy of the PC scores.
Alternative computational methods have been developed to improve the efficiency of PCA.
In this study, we compare four such methods: TRACE, projection, online TRACE, and HDPCA projection.

\subsubsection{TRACE}

TRACE implements the standard PCA method, with additional procedures for handling minor issues such as missing data.
An important improvement in TRACE on the standard method is the inclusion of Procrustes analysis.
Given two data sets with the same sample size and possibly different dimensions,
Procrustes analysis finds the linear transformation that maps one data matrix as close as possible to the other data matrix.
In our contex, the PC scores of the reference matrix is computed and we obtain the $n \times n$ matrix $V$.
For a study individal with genotype $y$, we decompose $\tilde{X} = (X, y)$ into $\tilde{X} = \tilde{U}\tilde{D}\tilde{V}^T$.
(TRACE uses eigendecomposition to find the ``$V$'' matrix to reduce computational complexity since we do not need the ``$U$'' matrix here.)
Let
\[
  V =
  \begin{bmatrix}
    V_{\text{ref}} & W
  \end{bmatrix},
  \quad
  \tilde{V} =
  \begin{bmatrix}
    \tilde{V}_{\text{ref}} & \tilde{W}_\text{ref} \\
    \tilde{V}_\text{stu} & \tilde{W}_\text{stu}
  \end{bmatrix},
\]
where the dimensions of $V_\text{ref}$, $\tilde{V}_\text{ref}$, and $\tilde{V}_\text{stu}$ are $n \times k_1$, $n \times k_2$, and $1 \times k_2$, respectively, with $ 1 \leq k_1 \leq k_2 \leq n+1$. 
We use Procrustes analysis to find
\[
  A, b = \argmin{A, b} d(V_\text{ref}, \tilde{V}_{\text{ref}}A+b)
\]
for some metric $d$.
Finally, we get the Procrustes-adjusted top $k_1$ PC scores for the study individual $V_\text{pro} = \tilde{V}_\text{stu} A + b$.
The algorithm is summarized as follows:
\begin{enumerate}
\item Compute $X^T X$.
  (Computational complexity: $\bO[n^2p]$.)  
\item Apply eigendecomposition to $X^T X$ to get $X^T X = V D^2 V^T$ (columns of $D$ and $V$ are sorted by PC variance with the first column corresponding to the greatest PC variance).
  ($\bO[n^3]$.)
\item Append $X^T y$ and its transpose to the last column and row of $X^T X$, respectively.
  The entry of the bottom-right corner is $y^T y$.
  This is exactly the same as computing $\tilde{X}^T \tilde{X}$,
  where $\tilde{X} = (X, y)$ is the reference group with the testing individual added.
  ($\bO[np]$.)
\item Apply eigendecomposition to $\tilde{X}^T \tilde{X}$ to get $\tilde{X}^T \tilde{X} = \tilde{V} \tilde{D}^2 \tilde{V}^T$.
  ($\bO[n^3]$.)
\item Project the first $k_2$ columns of $\tilde{V}$ to the first $k_1$ columns of $V$ by using Procrustes analysis to get the transformed PC scores of the testing individual (the last row of $V$),
  where $k_1 < k_2 \ll n$.
  (Computational complexity is ignorable if $k_2$ is not large compared to $n$.)
  \item Go to step 3 for the next testing individual.
\end{enumerate}

Total Computational complexity: $\bO[n^2p + m(np + n^3)]$.

TRACE is the slowest among the computational methods we study.
However, its result is also the most reliable.

\subsubsection{Projection}
An exchange for runtimes with accuracy is the projection method.
For this method, we simply find the PC's of the reference matrix and project the study individual's genotype onto them,
as shown below:

\begin{enumerate}
\item Compute $X^T X$.
  (Computational complexity: $\bO[n^2p]$.)  
\item Apply eigendecomposition to $X^T X$ to get $X^T X = V D^2 V^T$.
  ($\bO[n^3]$.)
\item Calculate $W = Y^T (X V D_1^{-1})$, where $Y$ is the data matrix for all the testing individuals, $D_1$ the first $k_1$ columns of $D$. ($\bO[mpk_1 + pnk_1]$.)
\end{enumerate}

Total computational complexity: $\bO[n^2p + n^3 + mpk_1 + pnk_1)] \approx \bO[n^2p + mk_1p)]$,
given that $k_1 \ll n \ll p$.

As the reference sample size increases,
the computational complexity of the projection method only increases quadratically,
(except for decomposing the reference matrix, which is done once for all study individuals,)
while that of TRACE increases cubically.
However, it is known that the projection method tends to underestimates the magnitude of the PC scores when $n \ll p$,
which is known as the ``shrinking bias''.

\subsubsection{Online TRACE}

A remedy for TRACE is to reduce the computational complexity withou much sacrifice of the accuracy.
This is achieved by the online TRACE method.
Notice that in TRACE, only the top $k_2$ PC's are used.
Online TRACE calculates a close approximation of the first $k_3$ columns of $\tilde{V}$ for some $k_2 \leq k_3 \ll n$ and keeps everything else the same as in TRACE, including Procrustes analysis. 
The algorithm is
\begin{enumerate}
\item Compute $X^T X$.
  (Computational complexity: $\bO[n^2p]$.)  
\item Apply eigendecomposition to $X^T X$ to get $X^T X = V D^2 V^T$.
  ($\bO[n^3]$.)
\item Calculate $U_3 = X V_3 D_3^{-1}$,
  where $V_3$ is the first $k_3$ columns of $V$,
  and $D_3$ is the diagonal matrix of the first $k_3$ diagonal entries of $D$.
  ($\bO[k_3 n p]$.)
\item Calculate 
  \[
    L = U_3^T y \quad \text{and} \quad K = y^T H,
  \]
  where $H$ is the normalized  $y - U_3L$
  ($\bO[k_3 p]$.)
\item Calculate $Q^T Q$, where
  \[
    Q = 
    \begin{bmatrix}
      D_3 & L \\
      0 & K
    \end{bmatrix}.
  \]
  ($\bO[k_3^3]$.)
\item Apply eigendecomosition to $Q^T Q$ to get $Q^T Q = \ddot{V}_3 \ddot{D}^2_3 \ddot{V}^T_3$.
  ($\bO[k_3^3]$.)
\item Calculate
  \[
    \tilde{V}_3 =
    \begin{bmatrix}
      V_3 & 0 \\
      0 & 1
    \end{bmatrix}
    \ddot{V}_3.
  \]
  ($\bO[nk_3^2]$.)
\item Project the first $k_2$ columns of $\tilde{V}_3$ to the first $k_1$ columns of $V$ by using Procrustes analysis to get the transformed PC scores of the testing individual,
  where $k_1 < k_2  < k_3 \ll n$.
  (Computational complexity is ignorable if $k_2$ is not large compared to $n$.)
  \item Go to step 4 for the next testing individual.
\end{enumerate}

Total computational complexity: $\bO[n^2p + n^3 + k_3np + m(k_3p + k_3^3 + nk_3^2]) \approx \bO[n^2p + m(k_3p + k_3^2n)]$, provided $k_3 \ll n \ll p$.

Here we see that the computational complexity of online TRACE for the study individuals increases quadratically with respect to the reference sample size.
This makes the runtimes of online TRACE close to those of the projection method.
The closeness between the results given by online TRACE and TRACE will be empirically shown later.

\subsubsection{HDPCA Projection}

As online TRACE reduces the computational complexity of TRACE,
we have HDPCA projection that improves the accuracy of projection without increasing the runtimes.
HDPCA reads the PC scores and variances calculated by projection and estimate the shrinkage factor to restore the unshrinked scores. 
The algorithm is
\begin{enumerate}
\item Compute $X^T X$.
  (Computational complexity: $\bO[n^2p]$.)  
\item Apply eigendecomposition to $X^T X$ to get $X^T X = V D^2 V^T$.
  ($\bO[n^3]$.)
\item Calculate $W = (Y^T X) V D^{-1}$, where $Y$ is the data matrix for all the testing individuals. ($\bO[mpn + mn^2]$.)
  \item Input $D$ and $W$ to the pc\_adjust function in the HDPCA function to obtain the PC scores adjusted for the shrinking bias. ($\bO[p]$ for finding the number of distant spikes (function select.nspike), $\bO[n]$ for adjusting for the shrinking bias (function hdpc\_est))
\end{enumerate}

Total computational complexity: $\bO[n^2p + mn(p + n) + p + n] \approx \bO[n^2p + mn(p+n)]$,
given that $m$ or $n$ is large.

The computational complexities between projection and HDPCA projection are the same.
The accuracy of HDPCA projection will be empirically shown later.

\subsubsection{Computational Complexity Summary}
A comparison of computatinal complexities of the four PCA methods are shown in \Cref{tbl:cplx}.

\begin{table} 
  \centering
  \begin{tabular}{|l|l|}
    \hline
    Method & Computational Complexity \\ 
    \hline
    TRACE & $\bO[n^2 p + mn(p + n^2)]$ \\
    \hline
    Projection & $\bO[n^2p + mk_1p)]$ \\
    \hline
    Online TRACE & $\bO[n^2 p + mk_3(p + k_3 n)]$ \\
    \hline
    HDPCA Projection & $\bO[n^2p + mn(p+n)]$ \\
    \hline
  \end{tabular}
  \caption{
    Comparison of computational complexity for the three methods.
    Here $p$ is the number of SNPs,
    $n$ is the number of individuals in the reference group,
    and $m$ is the number of individuals in the  testing group.
    For projection, $k_1$ is the number of PCs needed
    For Online TRACE, $k_3$ is the number of PCs calculated by online SVD, which is usually much less than $n$.}
  \label{tbl:cplx}
\end{table}

\subsection{Data}

In this study, we used both simulated data and real-world data.

\subsubsection{Simulation Studies}

We used the GGS software to simulate individuals migrating on a grid.
The parameters we chose to use are as follows:
\begin{enumerate}
\item L: Number of loci in each genealogy. (Fixed to $1000$.)
\item G: Number of genealogies. (Equal to $p / L$)
\item K: Side length of the grid. (Fixed to $2$.)
\item c: Number of haploids at each cell of the grid. (Equal to $2n / K^2$ and
$2m / K^2$)
\item M: Migration rate. (Varies)
\end{enumerate}

\subsubsection{Real-World Datasets}

The real-world datasets used in our study are the HGDP and 1000 Genomes datasets.
They contain the genotype readings of individuals from different ethnic groups over the world.
For better comparison, we select only the SNPs that are included in both datasets,
the number of which is approximately $6 \times 10^5$.
Moreover, we removed individuals who have at least parent also in the dataset.
The samples size is approximately $900$ for HGDP and $2300$ for 1000 Genomes.

\section{Results (Simulation Study)}

We now present the results of our study.

\subsection{Parameter Setting}

\begin{enumerate}
\item $n$ fixed, $m$ changing
\begin{enumerate}
\item $p = 100000$
\item $n = 600$
\item $m$: $\min = 1000$, $\max = 5000$, $\text{increment} = 500$ 
\item $migration = 100$
\end{enumerate}
\item $m$ fixed, $n$ changing
\begin{enumerate}
\item $p = 100000$
\item $n$: $\min = 1000$, $\max = 5000$, $\text{increment} = 500$ 
\item $m = 200$
\item $migration = 100$
\end{enumerate}
\end{enumerate}

\subsection{Computation Time}

\includegraphics[width=0.98\textwidth]{runtimes}

\subsection{Accuracy}

Next, we show the accuracy of each method.
We have two ways to measure the accuracy.
First, we can use the information from the simulation given by GGS to categorize the reference individuals into $K^2$ clusters and find the mean of each cluster,
which will be treated as the center of that cluster.
We then find the Euclidean distance between the study individuals and the centers,
again using the true locations in the grid to categorize the study individuals.

Second, we can use Online TRACE as the standard and find the Euclidean distance between the study individuals' PC scores calculated by Online TRACE and another method.

The parameter setting is the same as in comparing the runtimes.

\includegraphics[width=0.98\textwidth]{err_refcenter.pdf}

\includegraphics[width=0.98\textwidth]{err_online.pdf}

\section{Discussion}

\newpage

\begin{thebibliography}{}

\bibitem[Brand(2002)]{brand} Brand, M. (2002). Incremental singular value decomposition of uncertain data with missing values. {\it Computer Vision--ECCV 2002}, 707-720.

\bibitem[Dey \& Lee(2016)]{dey} Dey, R., \& Lee, S. (2016). Asymptotic properties of Principal Component Analysis and shrinkage-bias adjustment under the Generalized Spiked Population model. {\it arXiv preprint arXiv:1607.08647}.

\bibitem[Jolliffe(2002)]{jolliffe} Jolliffe, I. (2002). {\it Principal component analysis}. John Wiley \& Sons, Ltd.
Chicago

 \bibitem[Wang et al.(2015)]{wang} Wang, C., Zhan, X., Liang, L., Abecasis, G. R., \& Lin, X. (2015). Improved ancestry estimation for both genotyping and sequencing data using projection procrustes analysis and genotype imputation. {\it The American Journal of Human Genetics, 96}(6), 926-937.

\end{thebibliography}

    % \begin{enumerate}
    %     \item $X$ ($p$ by $n$): 
    %         Genotype readings for the reference group.
    %     \item $Y = (y_1 \cdots y_m)$ ($p$ by $m$): 
    %         Genotype readings for the study group.
    % \end{enumerate}

% \begin{enumerate}
% \item HGDP
%   The HGDP dataset contains individuals from different ethnic groups in the world. It includes about $6 \times 10^5$ SNPs and 
%   \begin{enumerate}
%   \item Individuals from different countries in the world.
%   \item We only use the SNPs that are also included in the 1000
%     Genomes dataset.
%   \item $p \approx 6 \times 10^5, n \approx 900$.
%   \end{enumerate}
% \item 1000 Genomes
%   \begin{enumerate}
%   \item Individuals from different countries in the world.
%   \item We only use the SNPs that are also included in the HGDP
%     dataset.
%   \item $p \approx 6 \times 10^5, n \approx 2300$.
%   \end{enumerate}
% \item Data filtering
%   \begin{enumerate}
%   \item We only used the SNPs that are included in both the HGDP and
%     1000 Genomes datasets.
%   \item We have removed the individuals with at least one parent in
%     the dataset.
%   \end{enumerate}
% \end{enumerate}

\end{document}